---
title: "Homework 4"
output: html_notebook
---
Izzi Grasso
Due: 4/13/2020

##Problem 1

####Create df
```{r}
df <- data.frame(
  Method = c(rep(x = "MethodA", times = 6),
             rep(x = "MethodB", times = 7),
             rep(x = "MethodC", times = 5)),
  Value = c(77, 54, 67, 74, 71, 66,
            60, 41, 59, 65, 62, 64, 52, 
            49, 52, 69, 47, 56)
)
```

Null hypothesis: Distributions are identical. 
Alt: Not null. 

$\alpha$ = 0.05. 

####Kruskal Wallis
```{r}
kruskal.test(Value ~ Method, data = df)
```
Since p < 0.05, we reject the null hypothesis at $\alpha$ = 0.05 and conclude there is a difference in efficacy of corrosion reduction methods. 

####Post-hoc
```{r}
pairwise.wilcox.test(df$Value, df$Method,
                 p.adjust.method = "bonferroni")
```
We can conclude Method A and Method B are significantly different. 

##Problem 2
####Data
```{r}

  Polluted <- c(21.3, 18.7, 23.0, 17.1, 16.8, 20.9, 19.7)
  Unpolluted  <-  c(14.2, 18.3, 17.2, 18.4, 20.0)

```

####Test for homogeneity
Null hypothesis: $\sigma_1^2$ = $\sigma_2^2$
Alternative hyp: $\sigma_1^2$ $\neq$ $\sigma_2^2$
```{r}
var.test(Polluted, Unpolluted)
```
At $\alpha$ = 0.05, we fail to reject null hypothesis. 


####Check for normality (Shapiro)
Null Hypothesis: normally distributed
Alternative Hyp: Not normally distributed
```{r}
shapiro.test(Polluted)
shapiro.test(Unpolluted)
```
Fail to reject null hypothesis. Assumption of normality holds. 

####Pooled t-test
Null Hypothesis: $\mu_1$ = $\mu_2$
Althernative Hyp: $\mu_1$ > $\mu_2$
```{r}
t.test(Polluted, Unpolluted, alternative = "greater", var.equal = TRUE)
```

We fail to reject the null hypothesis at $\alpha$ = 0.05. There is not significant evidence that the true fluoride concentration for livestock grazing is higher in the polluted region than the unpolluted region. 

##Problem 3
####Data 
```{r}
Day1 <- c(5.0, 4.8, 5.1, 5.1, 4.8, 5.1, 4.8, 4.8, 5.0, 5.2, 4.9, 4.9, 5.0)
Day2 <- c(5.8, 4.7, 4.7, 4.9, 5.1, 4.9, 5.4, 5.3, 5.3, 4.8, 5.7, 5.1, 5.7)
```


####Test for homogeneity
Null Hypothesis: $\sigma_1^2$ = $\sigma_2^2$
Alternative Hyp: $\sigma_1^2$ < $\sigma_2^2$
```{r}
var.test(Day1, Day2, alternative = "less", conf.level = .99)
```

At $\alpha$ = 0.01, we reject the null hypothesis. There is significant evidence that the variability of the process is greater on the second day than on the first. 

##Problem 4
####Data
```{r}
Men <- c(5, 10, 2, 0, 6, 4, 5, 15)
Women <- c(8, 9, 3, 5, 0, 4, 15)
```

####Test for normality (shapiro)
Null Hypothesis: normally distributed
Alternative Hyp: Not normally distributed
```{r}
shapiro.test(Men)
shapiro.test(Women)
```
Fail to reject null hypothesis. Assumption of normality holds. 

####Test for homogeneity
Null Hypothesis: $\sigma_1^2$ = $\sigma_2^2$
Alternative Hyp: $\sigma_1^2$ $\neq$ $\sigma_2^2$
```{r}
var.test(Men, Women)
```
Fail to reject null hypothesis. We can assume homogeneity. 

####Two sample t-test
```{r}
t.test(Men, Women, alternative = "two.sided", paired = FALSE, var.equal = TRUE, conf.level = 0.95)
```
At $\alpha$ = 0.05, we fail to reject the null hypothesis. There is not significant evidence that there is a difference in sick days between men and women. 

##Problem 5
####Data frame
```{r}
df3 <- data.frame(
  F1 = c(58, 29, 37, 40, 44, 37, 49, 49, 38),
  F2 = c(68, 67, 69, 58, 62, 48, 62, 76, 66),
  F3 = c(96, 90, 90, 103, 100, 91, 100, 114, 94),
  F4 = c(101, 110, 90, 103, 100, 91, 100, 114, 94),
  F5 = c(124, 114, 111, 113, 114, 102, 114, 112, 103)
)
```

####Test for normality
```{r}
shapiro.test(df3$F1)
shapiro.test(df3$F2)
shapiro.test(df3$F3)
shapiro.test(df3$F4)
shapiro.test(df3$F5)
```
Fail to reject null hypothesis. Normality assumption holds. 


####ANOVA
Null hypothesis: $\mu_1$ = $\mu_2$ = .. = $\mu_5$
Alternative hyp: Not null hypothesis
```{r}
df4 <- df3 %>% gather(Fertilizer, Value) 
model <- aov(Value ~ Fertilizer, df4)
summary(model)
```
At $\alpha$ = 0.01, we reject the null hypothesis. There is evidence that there is difference in yield by fertilizer. 


####Residual analysis
```{r}
plot(model)
```
Normality assumption holds, and homogeneity assumption holds. 

####Tukey test
```{r}
plot(TukeyHSD(model))
```
All yields are different from others except F4-F3. F5 produces the largest yield. 

####Kruskal Wallis
Null hypothesis: $\mu_1$ = $\mu_2$ = .. = $\mu_5$
Alternative hyp: Not null hypothesis
```{r}
kruskal.test(Value ~ as.factor(Fertilizer), data = df4)
```
At $\alpha$ = 0.01, we reject the null hypothesis. There is evidence that there is a difference of yield by fertilizer type. 

####Post hoc
```{r}
pairwise.wilcox.test(df4$Value, df4$Fertilizer,
                 p.adjust.method = "bonferroni")
```
At $\alpha$ = 0.01, F3-F4, F4-F5, F3-F5, and F1-F2 do not have a significant difference. 

##Problem 6
####Data 
```{r}
x <- c(47, 62, 65, 70, 75, 78, 95, 100, 120, 118)
y <- c(38, 62, 60, 67, 84, 79, 93, 104, 117, 150)

df5 <- data.frame(
  X = x,
  Y = y
)
```

####Plot
```{r}
ggplot(df5, aes(x = X, y = Y)) + 
  geom_point() +
  theme_linedraw()
```
Data looks linear, correlation analysis is reasonable. 

####Spearman Rank
Null hypothesis: $\rho$ = 0
Alternative hyp: $\rho$ $\neq$ 0
```{r}
cor.test(x, y, method = "spearman")
```

At $\alpha$ = 0.05 we reject the null hypothesis. There is a strong positive correlation between x and y. 

