---
title: "Homework 2"
output: html_notebook
---
Isabella Grasso
2/14/2020

##1. 
```{r}
library(phia)
library(gplots)

df <- data.frame(FactorS = as.factor(c(1,1,1,1,1,1,2,2,2,2,2,2,3,3,3,3,3,3)),
                 FactorF = as.factor(c(1,1,1,2,2,2,1,1,1,2,2,2,1,1,1,2,2,2)),
                 Value = c(189.7, 188.6,190.1,165.1,165.9,167.6,
                           185.1,179.4,177.3,161.7,159.8,161.6,
                           189.0,193.0,191.0,163.3,166.6,170.3)
                 )
```
####a. 
```{r}
plotmeans(Value ~ FactorF, data = df)
plotmeans(Value ~ FactorS, data = df)
```

####b. 
```{r}
model = lm(Value ~ FactorS + FactorF + FactorS:FactorF, data = df)
IM = interactionMeans(model)
plot(IM)
```
```{r}
anova(model)

```

Both factors are significant alone, but the interaction is insignificant at alpha = 0.05. 

##3. 

#### Create df

```{r}
df2 <- data.frame(FactorF = as.factor(c(1, 1, 1, 1, 1, 1, 1, 1,
                              2, 2, 2, 2, 2, 2, 2, 2)),
                  FactorT = as.factor(c(1, 1, 1, 1, 2, 2, 2, 2,
                             1, 1, 1, 1, 2, 2, 2, 2)),
                  Value = c(40, 24, 31, 21, 46, 53, 48,
                            46, 47, 37, 41, 41, 84, 76, 
                            77, 79)
                  )
```

#### ANOVA table
```{r}
model2 = lm(Value ~ FactorF + FactorT + FactorT:FactorF, data = df2)
anova(model2)
```

###Main effects & Interaction plot
```{r}
IM2 = interactionMeans(model2)
plot(IM2)

```
At $\alpha$ = 0.05, individual factors and interactions are all significant, and we reject the null hypothesis. There is a significant difference in yield with fertilizer 1 vs fertilizer 2, as well as with temperature and a significant interaction between temperature and fertilizer in crop yield. 


#### 6. 
```{r}
m <- matrix(c(42,309,31,64,246,46,20,115,27),ncol=3,byrow=TRUE)
colnames(m) <- c("We know","Cover up","Unsure")
rownames(m) <- c("Democrats","Republicans","Independents")
t <- as.table(m)
chisq.test(t)

```
At $\alpha$ = 0.05, we reject the null hypothesis and conclude that there is a difference in opinions by political party. 
