---
title: "Exam 3"
output:
  word_document: default
  html_notebook: default
---

Izzi Grasso
Due: 5/2/2020

```{r}
library(tidyverse)
library(ggplot2)
```


##Problem 1

####Data
```{r}
df1 <- data.frame(
  Depletion = c(85, 55, 40, 77),
  ModifiedHess = c(75, 45, 35, 67),
  Surber = c(31, 20, 9, 37),
  SubstrateRemoval = c(43, 21, 15, 27),
  Kicknet = c(17, 10, 8, 15)
)


```
####a. Plot
```{r}
df2 <- df1 %>%
  summarise_each(funs(mean)) %>%
  gather(Var, Val)

ggplot(df2, aes(x = as.factor(Var), y = Val)) + 
  geom_line(aes(group = 1)) +
  geom_point() + 
  theme_linedraw() + 
  labs( title = "Main Effects Plot",
        x = "Variables",
        y = "Mean Value")
  
```

Interpretation: Since the line is not at all flat, I would expect to reject the null hypothesis. Kicknet seems to be much lower than Modified Hess and Depletion.


####b. 
Null Hypothesis: The distributions are the same location
Alternative Hyp: Not null. 
```{r}
df1 <- df1 %>% 
  gather(Var, Val)

kruskal.test(Val ~ as.factor(Var), data = df1)
```

At $\alpha$ = 0.05 we reject the null hypothesis. There is significant evidence that the sampling procedures yield different species counts. 

####c. 
```{r}
pairwise.wilcox.test(df1$Val, df1$Var, p.adjust.method = "bonferroni")
```

At $\alpha$ = 0.05, we fail to reject the null hypothesis. 

##Problem 2 

####Data
```{r}
TarContent <- c(14, 17, 28, 17, 16, 13, 24, 25, 18, 31)
NicotineContent <- c(0.9, 1.1, 1.6, 1.3, 1.0, 0.8, 1.5, 1.4, 1.2, 2.0)
```

####a & b 
Null hypothesis: $\rho$ = 0
Alternative hyp: $\rho$ $\neq$ 0
```{r}
cor.test(TarContent, NicotineContent, method = "spearman")
```
 At $\alpha$ = 0.05 we reject the null hypothesis. There is a linear correlation between the amount of tar and the amount of nicotine found in cigarettes. 
 
##Problem 3

####Data 
```{r}
x = 261
n = 302
y = 401
m = 454
p_A = x/n
p_B = y/m

```

####a. 
```{r}
z_ = qnorm(0.005, mean = 0, sd = 1, lower.tail = FALSE)

LCI = (p_A - p_B) - z_*sqrt(x*(n-x)/(n^3)+y*(m-y)/(m^3))
UCI = (p_A - p_B) + z_*sqrt(x*(n-x)/(n^3)+y*(m-y)/(m^3))

LCI
UCI
```

####b.
```{r}
z_ = qnorm(0.05, mean = 0, sd = 1, lower.tail = FALSE)

LCI = (p_A - p_B) - z_*sqrt(x*(n-x)/(n^3)+y*(m-y)/(m^3))
UCI = (p_A - p_B) + z_*sqrt(x*(n-x)/(n^3)+y*(m-y)/(m^3))

LCI
UCI
```

####c.
```{r}
z_ = qnorm(0.025, mean = 0, sd = 1, lower.tail = FALSE)

LCI = (p_A - p_B) - z_*sqrt(x*(n-x)/(n^3)+y*(m-y)/(m^3))
UCI = (p_A - p_B) + z_*sqrt(x*(n-x)/(n^3)+y*(m-y)/(m^3))

LCI
UCI
```

####d. 
Null Hypothesis: $p_A$ = $p_B$
Alternative Hyp: $p_A$ $\neq$ $p_B$
```{r}
p_hat = (x+y)/(n+m)
z_star = (p_A-p_B)/sqrt(p_hat*(1-p_hat)*(1/n+1/m))
2*pnorm(z_star, mean = 0, sd = 1)

```

At $\alpha$ = 0.05 we fail to reject the null hypothesis. 


##Problem 4
Null Hypothesis: $s^2$ $\leq$ $\sigma^2$
Alternative hyp: $s^2$ > $\sigma^2$
```{r}

SSquared = 0.0002
SigmaSquared = 0.0003
alpha = 0.05
n = 10

TestStatistic <- ((n-1)*SSquared)/SigmaSquared

ChiSq <- qchisq((1-alpha), df = (n-1))

TestStatistic > ChiSq
```

##Problem 5

####Data
```{r}
df3 <- data.frame(
  Value = c(12.8, 11.7, 11.5, 12.6,
            10.6, 14.2, 14.7, 16.5,
            11.7, 11.8, 13.6, 15.4,
            10.7, 9.9, 10.7, 9.6,
            11.0, 13.8, 15.9, 17.1),
  Treatment = rep(1:4, times = 5),
  Block = rep(1:5, times = 1, each = 4)
)
```

####Friedman Test, RBD
Null hypothesis: Distributions are identical
Alternative hypothesis: Not null hypothesis. 
```{r}
friedman.test(df3$Value, df3$Treatment, df3$Block)
```

At $\alpha$ = 0.05, we fail to reject the null hypothesis. There is not significant evidence to show that there is a difference between treatment means. 